### 2019-09-10 delegations.tex
Change wording

\section{Quantum Delegation Schemes}

Here we describe schemes that allows a classical client to interact with a quantum server, and various properties that such schemes can have.

\begin{definition}
	A \emph{quantum delegation protocol} is an polynomial-round interactive protocol between a BPP verifier and a BQP prover. The BPP verifier has an arbitrary quantum circuit with classical input that it needs to evaluate by interacting with the prover.
\end{definition}

TODO (maybe don't assume universality?)

\begin{definition}
	A quantum delegation protocol has \emph{long output with error $\epsilon$} if the output can have multiple bits. In this case, the verifier should end with a measurement result of the output register. Said result's probability distribution would be within $\epsilon$ of the true distribution.
\end{definition}

\begin{definition}
	A quantum delegation protocol is \emph{secure} if a malicious prover cannot learn the input or the output of the computation. In particular, security notions such as \emph{CPA} can be applied here. (TODO Just define CPA right away since there are no other security notions around here)
\end{definition}

\begin{definition}
	A quantum delegation protocol is \emph{verifiable} if the verifier can check if the output is correct. TODO need to be more rigorous here.
\end{definition}

In \cite{mahadev_delegation}, a verifiable scheme is proposed. In \cite{mahadev_qfhe}, a secure scheme is proposed. (TODO just make this CPA or whatever it actually is...) We propose a verifiable scheme with long output which can also be made secure.

TODO check where does \cite{1904.06320} fit in.

\subsection{Our contributions}

We start by constructing a verifiable long output scheme. Then, we encrypt the scheme as given in \cite{mahadev_delegation} to achieve security. TODO Does it actually work though? There are multiple steps in the scheme... We encrypting each step separately or what?