### 2020-01-21 parallel/para_main.tex

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts,amsthm}
\usepackage{braket}
\usepackage{authblk}
%\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true,allcolors=blue]{hyperref}
\usepackage[margin=1in]{geometry}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage{mdframed} 
\mdfdefinestyle{figstyle}{ %
  linecolor=black!7, %
  backgroundcolor=black!7, %
  innertopmargin=10pt, %
  innerleftmargin=\textwidth, %
  innerrightmargin=\textwidth, %
  innerbottommargin=5pt %
}

\usepackage{comment}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{fact}{Fact}
\newtheorem{observation}{Observation}
\newtheorem{remark}{Remark}
\newtheorem{problem}{Problem}

\input{macros.tex}

\title{Classical Verification of Quantum Computations with \\ Efficient Verifier}
\begin{document}

\author[1]{Nai-Hui Chia}
\author[2]{Kai-Min Chung}
\author[3]{Takashi Yamakawa\thanks{This work was done in part while the author was visiting Academia Sinica.}}
\affil[1]{Department of Computer Science, University of Texas at Austin}
\affil[2]{Institute of Information Science, Academia Sinica}
\affil[3]{NTT Secure Platform Laboratories}
%\author{Nai-Hui Chia \and
%Kai-Min Chung \and
%Takashi Yamakawa}
%\date{November 2019}

\maketitle
\begin{abstract}
In this paper, we extend the protocol of classical verification of quantum computations (CVQC) recently proposed by Mahadev to make the verification efficient.
Our result is obtained in the following three steps:
\begin{itemize}
    \item We show that parallel repetition of Mahadev's protocol has negligible soundness error. This gives the first constant round CVQC protocol with negligible soundness error. In this part, we only assume the quantum hardness of the learning with error (LWE) problem similarly to the Mahadev's work.
    \item We construct a two-round CVQC protocol in the quantum random oracle model (QROM) where a cryptographic hash function is idealized to be a random function.
    This is obtained by applying the Fiat-Shamir transform to the parallel repetition version of the Mahadev's protocol.
    \item We construct a two-round CVQC protocol with efficient verifier in the CRS+QRO model where both prover and verifier can access to a (classical) common reference string generated by a trusted third party in addition to quantum access to QRO.
    Specifically, the verifier can verify a $\QTIME(T)$ computation in time $\poly(\secpar,\log T)$ where $\secpar$ is the security parameter.
    For proving soundness, we assume that a standard model instantiation of our two-round protocol with a concrete hash function (say, SHA-3) is sound and the existence of post-quantum indistinguishability obfuscation and post-quantum fully homomorphic encryption in addition to the quantum hardness of the LWE problem. 
\end{itemize}
\end{abstract}
\input{intro.tex}
\input{preliminaries.tex}
%\input{parallel.tex}
%\input{parallel_2.tex}
\input{parallel_3.tex}
\input{FiatShamir.tex}
\input{efficient_verification.tex}

\bibliographystyle{alpha}
\bibliography{abbrev3,crypto,reference}

\appendix
\input{appendix.tex}
\end{document}