### 2020-03-04 bqp-rewrite.tex
Add BQP Rewrite section

\section{REWRITE - Blind Delegation for BQP and FBQP}
\label{sec:BlindBQP2}

Let $\Pi$ be...
then there exists $\Pi$