### 2020-11-07 paper.tex

\newif\ifdraft
%\drafttrue
\draftfalse

\documentclass[11pt]{article}
\usepackage{authblk}
%\pagestyle{plain}

\usepackage{amsmath, amssymb, amsthm}
\usepackage{stmaryrd}
\usepackage{braket}
\usepackage{algorithm}
\usepackage{enumitem}
\usepackage{algpseudocode}
\usepackage[margin=1in]{geometry}
\newcommand{\sgn}{\operatorname{sgn}}
\DeclareMathOperator*{\E}{\mathbb{E}}
\DeclareMathOperator*{\spn}{\operatorname{span}}
\DeclareMathOperator*{\poly}{\operatorname{poly}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\usepackage{etoolbox}
\usepackage{hyperref}
\apptocmd{\thebibliography}{\raggedright}{}{}
\usepackage{cite,color,float}
\usepackage{mdframed} 
\usepackage{cleveref}
\crefname{protocol}{protocol}{protocols}
\Crefname{protocol}{Protocol}{Protocols}
\crefname{thm}{theorem}{theorems}
\Crefname{thm}{Theorem}{Theorems}
\crefname{rmk}{remark}{remarks}
\Crefname{rmk}{Remark}{Remarks}
\crefname{lem}{lemma}{lemmata}
\Crefname{lem}{Lemma}{Lemmata}

%*************************************page layout****************************
\iffalse
\setlength{\textheight}{9in}
\setlength{\columnsep}{2.0pc}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{0in}
\setlength{\headheight}{0ex}
\setlength{\hoffset}{0in}
\setlength{\headsep}{0.0in}
\setlength{\oddsidemargin}{0ex}
\setlength{\evensidemargin}{0ex}
\setlength{\parindent}{1pc}

\fi
%\pagecolor{black}
%\color{white}
\numberwithin{equation}{section}
\newcounter{protocol}
%\newcommand{\linefill}{\rule{\linewidth}{0.8pt}}

\newenvironment{protocol}[1]{\begingroup\setlength\parindent{0pt}\medskip\noindent\linefill\\
\refstepcounter{protocol}\textbf{Protocol \theprotocol} #1\\\noindent\linefill}
{\vspace{-\topsep}\noindent\linefill\endgroup}

\ifdraft
\newcommand{\KM}[1]{{\footnotesize\color{cyan}[KM: #1]}}
\newcommand{\Ethan}[1]{{\footnotesize\color{magenta}[Ethan: #1]}}
\newcommand{\XW}[1]{{\footnotesize\color{red}[XW: #1]}}
\else
\newcommand{\KM}[1]{}
\newcommand{\Ethan}[1]{}
\newcommand{\XW}[1]{}
\fi

\def \Hin {H_{\mathrm{in}}}
\def \Hout {H_{\mathrm{out}}}
\def \Hprop {H_{\mathrm{prop}}}
\def \Hclock {H_{\mathrm{clock}}}
\def \Jclock {J_{\mathrm{clock}}}
\def \Jprop {J_{\mathrm{prop}}}
\def \Kin {K_{\mathrm{in}}}
\def \Kclock {K_{\mathrm{clock}}}
\def \Kprop {K_{\mathrm{prop}}}

\newcommand{\histpsi}[1]{\ket{\psi_{#1}^{\mathrm{hist}}}}
\newcommand{\LHXZ}[1]{\mathrm{LH}_{\mathrm{XZ}}^{#1}}
\newcommand{\ground}[1]{{\lambda \left (#1 \right)}}

\title{Constant-round Blind Classical Verification of Quantum Sampling}

%\iffalse
\author[1]{Kai-Min Chung\thanks{\href{mailto:kmchung@iis.sinica.edu.tw}{kmchung@iis.sinica.edu.tw}. Partially supported by the 2019 Academia Sinica Career Development Award under Grant no. 23-17, and MOST QC project under Grant no. MOST 108-2627-E-002-001.}}
\author[2]{Yi Lee\thanks{\href{mailto:ethanlee515@gmail.com}{ethanlee515@gmail.com}. This work was done in part while the author was affiliated to National Taiwan University.}}
\author[3]{Han-Hsuan Lin\thanks{\href{mailto:linhh@cs.nthu.edu.tw}{linhh@cs.nthu.edu.tw}. Part of this work was done while supported by Scott Aaronson's Vannevar Bush Faculty Fellowship from the US Department of Defense.}}
\author[4]{Xiaodi Wu\thanks{\href{mailto:xwu@cs.umd.edu}{xwu@cs.umd.edu}. Partially supported by the U.S. National Science Foundation grant CCF-1755800, CCF-1816695, and CCF-1942837(CAREER).}}
\affil[1, 2]{Institute of Information Science, Academia Sinica, Taiwan}
\affil[3]{Department of Computer Science, National Tsing Hua University, Taiwan}
\affil[4]{
	Department of Computer Science, Institute for Advanced Computer Studies,
	and Joint Center for Quantum Information and Computer Science,
	University of Maryland, USA
}
%\fi
%\author{}
%\institute{}

\input{macros.tex}
\input{./parallel/macros.tex}
\begin{document}

\maketitle
%\thispagestyle{plain}

\begin{abstract}

In a recent breakthrough, Mahadev constructed a classical verification of quantum computation (CVQC) protocol for a  classical client to delegate decision problems in $\BQP$ to an untrusted quantum prover under computational assumptions. In this work, we explore further the feasibility of CVQC with the more general \emph{sampling} problems in BQP and with the desirable \emph{blindness} property. We contribute affirmative solutions to both as follows. 
\begin{itemize}
\item  Motivated by the sampling nature of many quantum applications (e.g., quantum algorithms for machine learning and quantum supremacy tasks), we initiate the study of  CVQC for \emph{quantum sampling problems} (denoted by $\SampBQP$).  More precisely, in a CVQC protocol for a $\SampBQP$ problem, the prover and the verifier are given an input $x\in \zo^n$ and a quantum circuit $C$, and the goal of the classical client is to learn a sample from the output $z \leftarrow C(x)$ up to a small error, from its interation with an untrusted prover. We demonstrate its feasibility by constructing a four-message CVQC protocol for $\SampBQP$ based on the quantum \emph{Learning With Error} assumption.

\item
The \emph{blindness} of CVQC protocols refers to a property of the protocol where the prover learns nothing, and hence is blind, about the client's input. It is a highly desirable property that has been intensively studied for the delegation of quantum computation. 
Somewhat surprisingly, we provide a simple yet powerful \emph{generic} compiler that transforms any CVQC protocol to a blind one while preserving the completeness and soundness errors.  
\end{itemize}
Applying our compiler to (a parallel repetition of) Mahadev's CVQC protocol for $\BQP$ and our CVQC protocol for $\SampBQP$ yields the first \emph{constant-round} blind CVQC protocol for $\BQP$ and $\SampBQP$ respectively, with negligible completeness and soundness errors.

\vspace{1mm}
\noindent \textbf{Keywords:} classical delegation of quantum computation, blind quantum computation, quantum sampling problems

\iffalse

\KM{version 2}

In a recent breakthrough, Mahadev constructed a classical verification of quantum computation (CVQC) protocol for a  classical client to delegate \emph{decision problems} in $\BQP$ to an untrusted quantum prover. In this work, motivated by the fact that many applications of quantum computations (e.g., in quantum machine learning) are sampling problems, we initiate the study of CVQC for \emph{quantum sampling problems} (denoted by $\SampBQP$), where given an input $x\in \zo^n$, the goal is to learn a sample an output $z \leftarrow C(x)$ over  $\zo^m$ for a quantum circuit $C$ from an untrusted prover with a correct distribution (up to a small error $\eps$).

As our main result, we construct a four-message CVQC protocol for $\SampBQP$ with an arbitrarily small inverse polynomial error $\eps$ based on the quantum learning with error assumption.

\fi

%Following the recent breakthrough of Mahadev, we explore the feasibility of classical verification of quantum computation (CVQC) for a broader class of quantum computation with blindness property.

%This is a placeholder. Lorem ipsum. Lorem ipsum. Lorem ipsum...

\end{abstract}

\newpage

\input{introduction}

\input{preliminaries}

\input{model}

\input{hamiltonian}

%\input{qpip}

\input{qpip0}

\input{blind}

\bibliographystyle{plain}
\bibliography{refs}

\end{document}