### 2021-05-31 hamiltonian_short.tex
Sync Overleaf updates with GitHub

\section{Construction of the $\QPIP_1$ Protocol for $\SampBQP$}
\label{sec:sampbqp_short}

See \Cref{sec:sampbqp} for the details of the construction of our $\QPIP_1$ Protocol for $\SampBQP$, \Cref{ProtoQPIP1}. The soundness and completeness of \Cref{ProtoQPIP1} is stated in the following theorem.

\begin{thm}
    \label{QPIP1thm}
	$\PiSamp$ is a $\QPIP_1$ protocol for the $\SampBQP$ problem\\ $(D_x)_{x\in\set{0,1}^*}$ with negligible completeness error and is statistically sound.\footnote{The soundness and completeness of a $\SampBQP$ protocol is defined in \Cref{dfn:stats-secure-proto-sampbqp}.}
	%\XW{do we want precise error bound?}
\end{thm}