### 2021-06-02 acknowledgements.tex
Sync for submission

\section*{Acknowledgments}

The authors would like to thank Tomoyuki Morimae for valuable feedback that helped improve the paper.
We are also thankful to the anonymous reviewers from STOC 2021 for various useful comments and mentioning the stronger security guarantees given by Brakerski's encryption scheme in \cite{brakerski_qfhe}. \Ethan{check if this is ok}