### 2021-06-03 acknowledgements.tex
Sync submitted version

\section*{Acknowledgments}

The authors would like to thank Tomoyuki Morimae for his valuable feedback that helped improve the paper and for pointing out the related works \cite{takeuchi2018verification, hayashi2015verifiable}.
We are also thankful to anonymous reviewers for various useful comments. \Ethan{check if this is ok}