### 2021-06-03 qpip0_proofs.tex

% \begin{theorem}[binding property of $\PiNaive$]
% 	\label{lem:naive-qpip0-binding}
% 	Let $\PNaiveStar$ be a cheating $\BQP$ prover for $\PiNaive$ and $\lambda$ be the security parameter.
% 	Suppose that $\Prob{d=\Acc\mid y\ne\bot, c=0}$ is overwhelming, 
% 	under the QLWE assumption, then the verifier's output in the Hadamard round is $O(\eps)$-computationally indistinguishable from $(d, z_{ideal})$.
% \end{theorem}